---
title: "DP First Laboratory"
output:
  word_document: default
  html_document:
    df_print: paged
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results="hide")
```

```{r load, message=FALSE, warning=FALSE}
library(sdcMicro)
```

```{r dataset}
# we can use the instruction data() to show the available datasets
data("free1") # loads the dataset
#we use <- to assign a value to a variable
newdataset<-free1 # newdataset is a copy of the free1 dataset
```

```{r dataframe}
newdataset<-as.data.frame(newdataset)
class(newdataset)
str(newdataset) #new structure therefore, new attributes
attributes(newdataset)
# ?data.frame  # you can use a question mark before a command to obtain some help
```

```{r}
names(newdataset)[1:4]
```
Checking the output of the previous commands we can see the first 4 variables of the dataset being: Region, Sex, Age, Marstat.
So that, Region, Sex and Marital Status are Categorical while Age would be continuous.

```{r}
newdataset$REGION<-as.factor(newdataset$REGION)
newdataset$SEX<-as.factor(newdataset$SEX)
newdataset$AGE<-as.factor(newdataset$AGE)
newdataset$MARSTAT<-as.factor(newdataset$MARSTAT)
```


```{r}
# Levels of the region parameter
r = levels(newdataset$REGION)
# Levels of the sex parameter
s = levels(newdataset$SEX)
# Levels of the age parameter
a = levels(newdataset$AGE)
# Levels of the marital status parameter
m = levels(newdataset$MARSTAT)
```

```{r}
length(r)*length(s)*length(a)*length(m)
```
Easy to disclure.

```{r}
contable = table(newdataset$REGION,newdataset$SEX)
```

```{r}
print(contable)
```

```{r}
sum(contable[contable < 2])
sum(contable[contable < 3])
```

```{r}
contable4 = table(newdataset$REGION,newdataset$SEX,newdataset$AGE,newdataset$MARSTAT)
```

```{r}
sum(contable4[contable4 < 2])
sum(contable4[contable4 < 3])
```

```{r}
freqCalc(newdataset,keyVars = c("REGION","SEX"))
```
```{r}
freqCalc(newdataset,keyVars = c("REGION","SEX","AGE","MARSTAT"))
```
```{r}
dataset32<-free1
```

```{r}
dataset32<-as.data.frame(dataset32)
```

```{r}
sdc <- createSdcObj(
  dat = dataset32,
  keyVars = c("REGION","SEX","AGE","MARSTAT")
)
sdc <- varToFactor(sdc, "REGION")
sdc <- varToFactor(sdc, "SEX")
sdc <- varToFactor(sdc, "AGE")
sdc <- varToFactor(sdc, "MARSTAT")
```

```{r}
print(sdc, type="risk")
print(sdc, type="kAnon")
```

```{r}
slotNames(sdc)
```

```{r}
sdc@risk
```

4 SDC methods:

4.1 Recoding:

First of all, we create a list with all the regions, and a list with the different values of the regions list defined. This two lists will be used later on to recode the regions values.
```{r}
regions <- sdc@manipKeyVars$REGION
diff_regions <- levels(regions)
```

Now, we construct a list with the different categories: North(45), South(45), Est(45) and West(total-135) following the instructions in the statement.

```{r}
groups <- c(rep("NORTH",45), rep("SOUTH",45), rep("EST",45), rep("WEST", length(diff_regions)-135))
```

Then we just need to use the groupAndRename function specifying the previous values, being the numbers from 1 to 182, and the new ones we want to convert to being the cardinal directions defined above. The groupAndRename function will swap each element from the before list for the value in the after list in the region column in the sdc object.

```{r}
nRegions <- groupAndRename(obj = sdc, var = c("REGION"), before = c(diff_regions), after = c(groups))
```

Now we can construct a table to count how many different cardinal points we have.

```{r}
table(nRegions@manipKeyVars$REGION)
```
Now, just to check that everything worked properly we do the following checks:
```{r}
checkCorrectness <- table(sdc@manipKeyVars$REGION)
sum(checkCorrectness[1:45])
sum(checkCorrectness[46:90])
sum(checkCorrectness[91:135])
sum(checkCorrectness[136:length(diff_regions)])
```
We can now check the values of 2/3 anonymity and see how they have been modified when applying our changes.

```{r}
print(nRegions, type="kAnon")
```

```{r}
sdc <- varToNumeric(obj = sdc, var = "AGE")
```

```{r}
intervals <- c(0,14,25,64,Inf)
```


```{r}
nAge <- globalRecode(obj = sdc, column = c("AGE"), intervals, labels = c("Children","Young","Adults","Senior"))
```

```{r}
table(nAge@manipKeyVars$AGE)
```

```{r}
table(sdc@manipKeyVars$AGE)
```

```{r}
print(nAge, type="kAnon")
```

4.2 Local Suppression:
```{r}
class(newdataset)
```

```{r}
#sdcApp()
```

4.2 a) b) c) d)

###

library(sdcMicro)

obj <- NULL
if (!exists("newdataset")) {
  stop('object "newdataset" is missing; make sure it exists.`', call. = FALSE)
}
obj$inputdata <- readMicrodata(path="newdataset", type="rdf", convertCharToFac=FALSE, drop_all_missings=FALSE)
inputdataB <- obj$inputdata

## Set up sdcMicro object
sdcObj <- createSdcObj(dat=inputdata,
	keyVars=c("REGION","SEX","AGE","MARSTAT"), 
	numVars=c("KINDPERS","NUMYOUNG","NUMOLD","AGEYOUNG","EDUC1","EDUC2","ETNI","PRIOCCU","POSLABM","REGJOBC","RECBEN","RECUNBEN","RECODBEN","RECBILL","RECSOSEC","RECPENS","POSLABLY","POSFACT","COMPCODE","OCCUCODE","KINDFACT","TENURE","FTPTIME","ADDJOB","JOBFIND","WEIGHT","INCOME","MONEY","ASSETS","DEBTS"), 
	weightVar=NULL, 
	hhId=NULL, 
	strataVar=NULL, 
	pramVars=NULL, 
	excludeVars=NULL, 
	seed=0, 
	randomizeRecords=FALSE, 
	alpha=c(1))

## Store name of uploaded file
opts <- get.sdcMicroObj(sdcObj, type="options")
opts$filename <- "newdataset"
sdcObj <- set.sdcMicroObj(sdcObj, type="options", input=list(opts))


## Local suppression to obtain k-anonymity
sdcObj <- kAnon(sdcObj, importance=c(4,1,3,2), combs=NULL, k=c(3))
sdcObj <- undolast(sdcObj)
## Local suppression to obtain k-anonymity
sdcObj <- kAnon(sdcObj, importance=c(4,1,3,2), combs=NULL, k=c(5))
sdcObj <- undolast(sdcObj)
## Local suppression to obtain k-anonymity
sdcObj <- kAnon(sdcObj, importance=c(1,4,2,3), combs=NULL, k=c(3))
sdcObj <- undolast(sdcObj)
## Local suppression to obtain k-anonymity
sdcObj <- kAnon(sdcObj, importance=c(1,4,2,3), combs=NULL, k=c(5))

###

4.2 e)

###

library(sdcMicro)

obj <- NULL
if (!exists("newdataset")) {
  stop('object "newdataset" is missing; make sure it exists.`', call. = FALSE)
}
obj$inputdata <- readMicrodata(path="newdataset", type="rdf", convertCharToFac=FALSE, drop_all_missings=FALSE)
inputdataB <- obj$inputdata

## Convert a numeric variable to factor (each distinct value becomes a factor level)
inputdata <- varToFactor(obj=inputdata, var=c("RECBEN"))
## Set up sdcMicro object
sdcObj <- createSdcObj(dat=inputdata,
	keyVars=c("SEX","AGE","MARSTAT","RECBEN"), 
	numVars=NULL, 
	weightVar=NULL, 
	hhId=NULL, 
	strataVar=NULL, 
	pramVars=NULL, 
	excludeVars=NULL, 
	seed=0, 
	randomizeRecords=FALSE, 
	alpha=c(1))

## Store name of uploaded file
opts <- get.sdcMicroObj(sdcObj, type="options")
opts$filename <- "newdataset"
sdcObj <- set.sdcMicroObj(sdcObj, type="options", input=list(opts))


## Adding linked (ghost)-Variables
sdcObj <- addGhostVars(sdcObj, keyVar="RECBEN", ghostVars=c("REGION"))
## Local suppression to obtain k-anonymity
sdcObj <- kAnon(sdcObj, importance=c(3,1,2,4), combs=NULL, k=c(10))
sdcObj <- undolast(sdcObj)
## Local suppression to obtain k-anonymity
sdcObj <- kAnon(sdcObj, importance=c(3,1,2,4), combs=NULL, k=c(10))

###

